---
title: "codebook"
author: "Marina T"
date: "April 25, 2015"
output: html_document
---

This code book describes variables, data, and transformations that where performed to clean up the following dataset: Human Activity Recognition Using Smartphones Dataset (Version 1.0).  It contains accelerometer measurements from Samsung Galaxy S smartphones while subjects performed various activities.


The following files are used from the CI HAR Dataset

* activity_labels
* features
* subject_test
* subject_train
* X_train
* X_test
* y_train
* y_test



The following names are used: Activity, Subject and Features, as part of descriptive variable names for data in data frame.

First step is to load data from the CI HAR dataset.

-Here is a sample of code:

```{r, echo=TRUE}
x_train <-read.table("~/Desktop/Coursera/Getting Cleaning Data/X_train.txt", header = FALSE)

x_test <-read.table("~/Desktop/Coursera/Getting Cleaning Data/X_test.txt", header = FALSE)
```

```{r}
y_train <-read.table("~/Desktop/Coursera/Getting Cleaning Data/y_train.txt", header = FALSE)
y_test <-read.table("~/Desktop/Coursera/Getting Cleaning Data/y_test.txt", header = FALSE)
activity_labels <-read.table("~/Desktop/Coursera/Getting Cleaning Data/activity_labels.txt")
subject_test <-read.table("~/Desktop/Coursera/Getting Cleaning Data/subject_test.txt", header = FALSE)
subject_train <-read.table("~/Desktop/Coursera/Getting Cleaning Data/subject_train.txt", header = FALSE)
```
**To merge all of the data into one data set, the following steps were done:**

* Consolidate by rows:

```{r, echo = TRUE}
data_subject <- rbind(subject_train, subject_test)
data_y <- rbind(y_train, y_test)
data_x <- rbind(x_train, x_test)
```

* Assign proper names to the variables

```{r, echo=TRUE}
names(data_subject) <- c("subject")
names(data_y) <- c("activity")

features <- read.table("~/Desktop/Coursera/Getting Cleaning Data/features.txt", head = FALSE)
names(data_x) <- features$V2
```

* Combining data, by subject and activity:

```{r, echo=TRUE}
bind_data <- cbind(data_subject, data_y)
data <- cbind(data_x, bind_data)
```

**The following code is calculating Mean and Standar Deviation**


```{r, echo=FALSE}
CalcData <- features$V2[grep("mean\\(\\)|std\\(\\)", features$V2) ]
```

```{r}
##subset of data for analysis
selectNames <- c(as.character(CalcData), "subject", "activity")
data <- subset (data, select = selectNames)
```
**Appropriate labels are assigned:**

```{r, echo=TRUE}
names(data)<-gsub("^t", "time", names(data))
names(data)<-gsub("^f", "frequency", names(data))
names(data)<-gsub("Acc", "Accelerometer", names(data))
names(data)<-gsub("Gyro", "Gyroscope", names(data))
names(data)<-gsub("Mag", "Magnitude", names(data))
names(data)<-gsub("BodyBody", "Body", names(data))
```
```{r}
## tidy output

library(plyr);
output <-aggregate(. ~subject + activity, data, mean)
output <- output[order(output$subject, output$activity), ]
```

**Finally, the output is written to a text file:**
```{r,echo=TRUE}
write.table(output, file = "tidydataset.txt", row.name = FALSE)
```



